---
title: "filter_join.Rmd"
author: "Annette Brickley"
date: "5/14/2021"
output: html_document
---

```{r setup, include=TRUE}
library(tidyverse)
library(readxl)
library(here) 
library(kableExtra)
```

## Read in data
```{r dataread}
fish <- read_csv(here("data", "fish.csv"))
kelp_abur <- read_excel(here("data", "kelp.xlsx"),
                        sheet = "abur")
```

## Filter rows by matching a single char string, == means exactly equal
```{r filter1}
fish_garibaldi <- fish %>% 
  filter(common_name == "garibaldi")
fish_mohk <- fish %>% 
  filter(site == "mohk")
```

## Filter rows based on numeric conditions
```{r filter2}
fish_over50 <- fish %>% 
  filter(total_count >= 50)
```

## Filter for a subset of the data
```{r filter3}
fish_3sp <- fish %>% 
  filter(common_name == "garibaldi" |  # | means OR
           common_name == "blacksmith" |
           common_name == "black surfperch")
# can do the same thing with code below because they are all in common_name
# filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))
```

### Filter for a subset based on either of two different columns
```{r filter3b}
fish_gar_2016 <- fish %>% 
  filter(year == 2016 | common_name == "garibaldi")
```

## Filter to for a subset based on both (or more) columns using a , or & for AND
```{r filter4}
aque_2018 <- fish %>% 
  filter(year == 2018 & site == "aque")
```

## Combining these tactics
```{r filter3,4}
low_gb_wr <- fish %>% 
  filter(common_name == "garibaldi" |
           common_name == "rock wrasse" &
           total_count <= 10)
```

## Looking for specific strings, can use str_detect and define a pattern
```{r filter5}
fish_bl <- fish %>% 
  filter(str_detect(common_name, pattern = "black"))
fish_nogar <- fish %>% 
  filter(str_detect(common_name, pattern = "gari", negate = TRUE)) # excludes any with pattern
```

## Create a subset that looks for certain values within specific columns
```{r filter6}
low_gb_wr <- fish %>% 
  filter(common_name %in% c("garibaldi", "rock wrasse"),
         total_count <= 10)
```

## Converting 2 digit year to 4 digit year, all 20xx (Annette's way)
```{r year2-4}
kelp_abur_4y <- kelp_abur %>%
 select(year, site, total_fronds) %>%
    mutate(year = year + 2000)
```

# Joining data frames
## full_join() keeps everything-- if in doubt, full join! Using kelp_abur as our “left” data frame, and fish as our “right” data frame
```{r join_full}
abur_kelp_fish <- kelp_abur_4y %>%  # chooses left data
  full_join(fish, by = c("year", "site")) # chooses right data and variables to merge by
```

## left_join will join the variables in fish that match kelp_abur
```{r join_to_left}
kelp_fish_left <- kelp_abur_4y %>% 
  left_join(fish, by = c("year", "site"))
```

## can also right_join and inner_join, but not doing another example here.

## combining skills in sequence
```{r filter-join-mutate}
my_fish_join <- fish %>% 
  filter(year == 2017, site == "abur") %>% 
  left_join(kelp_abur_4y, by = c("year", "site")) %>% 
  mutate(fish_per_frond = total_count / total_fronds)
```
# Formatting a table for document
## default Kable format
```{r}
kable(my_fish_join)
```

## fancier with kable_styling
```{r}
my_fish_join %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```
