---
title: "my.tidying.Rmd"
author: "Annette Brickley"
date: "5/13/2021"
output: html_document
---

## Attach packages
```{r setup, include=TRUE}
library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(skimr)
```

## Read in new data for cleaning and explore
```{r dataread}
inverts <- read_excel(here("data", "inverts.xlsx")
          )
skimr::skim(inverts)
```

## We’ll use tidyr::pivot_longer() to gather data from all years in inverts (columns 2016, 2017, and 2018) into two columns, switching from wide format to long format
```{r switchwide-long}
inverts_long <- pivot_longer(data = inverts,
                             cols = '2016':'2018',
                             names_to = "year",
                             values_to = "sp_count")
```

## But using class(), we can see that the years are still character strings and not numbers. We need to coerce that column by mutating it.
```{r coercing}
inverts_long <- inverts_long %>%
  mutate(year = as.numeric(year))
```

## If we wanted to split the common name column up into individual species columns, we can use pivot_wider to spread
```{r switchlong-wide}
inverts_wide <- inverts_long %>% 
  pivot_wider(names_from = common_name, 
                     values_from = sp_count)
```

## To clean up column names to the default lower_snake_case using janitor
```{r columnnameclean}
inverts_wide <- inverts_wide %>%
  clean_names()
```

## To merge two or more columns, use tidyr::unite 
```{r columncombine}
inverts_unite <- inverts_long %>%
  unite(col = "site_year", # what to name the new united column
        c(site, year), # the columns being united
        sep = "_") # how to separate the things we're uniting
```

## To separate two or more columns, use tidyr::separate
```{r columnsep}
inverts_sep <- inverts_unite %>%
  separate(site_year, # name of col to separate
          into = c("my_site", "my_year")) # what to name new cols
```

## To find and replace a string of characters with something else, use stringr::str_replace
```{r findrepl}
ca_abbr <- inverts %>%
  mutate(
    common_name = 
      str_replace(common_name,
                  pattern = "california",
                  replacement = "CA"))
```
