---
title: "pivot_lobsters"
author: "Annette Brickley"
date: "5/12/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
```

## read in data from Excel (knowing it's an xlsx file) and removing top 4 lines of metadata

```{r dataread}
lobsters <- read_xlsx(here("data/lobsters2.xlsx"),
            skip=4)
```

## explore data
```{r explore}
skimr::skim(lobsters)
```

## group data by one variable
```{r group1}
lobsters %>%
  group_by(year) %>%
  summarize(count_by_year = n())
```

## group data by two variables
```{r group2}
lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_siteyear = n())
```

## summarize multiple variables including a mean and sd. But remove NA's
```{r summarize and assign}
siteyear_summary <- lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_siteyear =  n(),
            mean_size_mm = mean(size_mm, na.rm=TRUE),
            sd_size_mm = sd(size_mm, na.rm=TRUE)
  )
```

## inspect our new variable
```{r}
siteyear_summary 
```

## Include a median with our new variable
```{r}
siteyear_summary <- lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_siteyear =  n(),
            mean_size_mm = mean(size_mm, na.rm=TRUE),
            sd_size_mm = sd(size_mm, na.rm=TRUE),
            median_size_mm = median(size_mm, na.rm = TRUE)
  )
```

## Make a graph with our new variable
```{r graph}
ggplot(data = siteyear_summary, 
       aes(x = year, y = median_size_mm, color = site)) +
       geom_line() +
       facet_wrap(~site)
```

## Saving and exporting the most recently run ggplot graph with added dimensions
```{r savinglastplot}
ggsave(here("figures", "medlobsite_graph.png"), dpi = 180, width = 5, height = 4)
```

## We can mutate or add columns with calculations or repeated numbers or character strings. Good to rename since modifying original data.
```{r addcolumns}
lobsters_detailed <- 
  lobsters %>% 
  mutate(size_m = size_mm / 1000,
         millenia = 2000,
         observer = "Allison Horst")
```

## We can select which lets us move, choose, retain columns into a smaller subset of orginal.
```{r movecolumns}
lobsters_cut <-
  lobsters_detailed %>%
     select(date, site, size_m)
```

