---
title: "Creating graphs with R in ggplot2"
author: "Annette Brickley"
date: "5/10/2021"
output: html_document
---

## Plots with ggplot2

We are going to make plots in R and it's going to be amazing.

## Attach packages to work with data, read in Excel file, and make plots
```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(here)
```

## Read in csv file from data folder and assign a name
```{r read_csv}
ca_np <- read_csv(here("data", "ca_np.csv"))
head(ca_np)
```

## Read in xls file from data folder and assign a name
```{r read_xls}
ci_np <- read_excel(here("data", "ci_np.xlsx"))
head(ci_np)
```

## Creating a bare-bones ggplot line graph
```{r graph}
ggplot(data = ci_np, aes(x = year, y = visitors)) + geom_line()
```

## Save that basic plot as an object to reuse for other plots
```{r basegraph}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))
```

## To customize graphs with aesthetics by adding arguments within layer where element was created
```{r custom1}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

## To find which colors are recognized, google "R colors ggplot2". Can also customize fill, pch, size, alpha and more.If aesthetic based on variable, include aes
```{r custom2}
gg_base +
  geom_col(aes(fill = visitors)
  )
```

