---
title: "Creating graphs with R in ggplot2"
author: "Annette Brickley"
date: "5/10/2021"
output: html_document
---

## Plots with ggplot2

We are going to make plots in R and it's going to be amazing.

## Attach packages to work with data, read in Excel file, and make plots
```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(here)
```

## Read in csv file from data folder and assign a name
```{r read_csv}
ca_np <- read_csv(here("data", "ca_np.csv"))
head(ca_np)
```

## Read in xls file from data folder and assign a name
```{r read_xls}
ci_np <- read_excel(here("data", "ci_np.xlsx"))
head(ci_np)
```

## Creating a bare-bones ggplot line graph
```{r graph}
ggplot(data = ci_np, aes(x = year, y = visitors)) + geom_line()
```

## Save that basic plot as an object to reuse for other plots
```{r basegraph}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))
```

## To customize graphs with aesthetics by adding arguments within layer where element was created
```{r custom1}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

## To find which colors are recognized, google "R colors ggplot2". Can also customize fill, pch, size, alpha and more.If aesthetic based on variable, include aes
```{r custom2}
gg_base +
  geom_col(aes(fill = visitors)
  )
```

## Can also apply themes to graphs that take care of many basic aesthetics.
```{r customtheme}
gg_base + 
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()
```

## Use labs() to update axis labels, titles
```{r labels}
gg_base +
  geom_col(fill = "cornflowerblue") +
  theme_bw() +
  labs(
    x = "Year",
    y = "Annual park visitors",
    title = "Channel Islands NP Visitation",
    subtitle = "(1963 - 2016)"
  )
```

## To combine two compatible graph types as layers, like adding points to a line graph.
```{r combining}
gg_base +
  geom_line(color = "purple") +
  geom_point(color = "orange",
             alpha = 0.5
  )
```

## To graph datasets with multiple series, need to group by name or color or other. Switching to ca_np dataset with multiple park names.
```{r multiseries}
ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)) + geom_line(aes(color = park_name))
gg_np <- ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)
  )
```

## Faceting graphs or splitting into multiple panels based on subsets of the data. Example split by park.
```{r faceting}
gg_np + 
  geom_line(show.legend = FALSE) +
  theme_light() +
  labs(x = "year", y = "annual visitors") +
  facet_wrap(~ park_name)
```

## Saving and exporting the most recently run ggplot graph with added dimensions
```{r savinglastplot}
ggsave(here("figures", "np_graph.jpg"), dpi = 180, width = 5, height = 4)
```
